opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-12-04 chapters/03_existing_datasets.tex
Commited table version with footnote directly underneath
Added labeled column in dataset overview

\begin{table}[!htb]
\centering
\caption{Overview of Experimental Powder Diffraction Databases}
\label{tab:exp_databases}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Name}                          & \textbf{No. of Patterns} & \textbf{Year Established} \\
\midrule
Powder Diffraction File (PDF)  \tablefootnote{The PDF lists the Materials Phases Data System (MPDS) as a database source. The MPDS and Pearson's crystal, both maintained by the Pauling File project, contain largely identical data. Hence it is likely that there is significant overlap in the experimental patterns available in the PDF and Pearson's crystal data.}         & 20,800                  & 1941                      \\
Pearson's Crystal Data                 & 21,700                  & 2007        \\
Crystallography Open Database (COD)    & 1063                    & 2003                      \\
RRUFF                                  & 1290                    & 2006                      \\
PowBase                                & 169                     & 1999                      \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Powder Diffraction File:} \cite{PDFWeb} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941\cite{GatesRector2019}. According to the ICDD the PDF5+, the latest release of the PDF as of November 2024, contains over a million materials with accompanying powder diffraction data. However, most of these powder diffraction patterns are simulated. After inquiring with the ICDD in April 2024 we were told that only 20,800 of the powder diffraction patterns in the PDF5+ stemmed from experiments. Of these 20,800 entries, 10,954 contain information about the atomic coordinates of the underlying structures.%\footnote{We reached out to the contact address given on their website, info@icdd.com.} 
Still, this makes the PDF5+, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers. As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00. However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published.\footnote{This information can be found in the product license agreement for the PDF5+ at \url{https://www.icdd.com/licensing-process/\#1528471154226-933e5cc6-8da7}.} \\

\textbf{Pearson's Crystal data}:\cite{PearsonWeb} Pearson's crystal data is a commercial database published and maintained by the Pauling File project \cite{kaduk2007}. The database, first published in 2007, currently contains 395,000 crystal structures and 21,700 corresponding experimental powder diffraction patterns. The Pauling File project also distributes the Material phases data system, which contains largely identical data. Since the PDF5+ lists the MPDS as a database source, there is likely significant overlap between the experimental patterns found in the PDF5+ and those found in Pearson's crystal data. As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}.\\

>>>>

\begin{table}[!htb]
\centering
\caption{Overview of Experimental Powder Diffraction Databases}
\label{tab:exp_databases}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Name}                          & \textbf{No. of Patterns} & \textbf{Year established} \\
\midrule
Linus Pauling file                 & 21,700                  & 2002        \\
Powder Diffraction File (PDF)  \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source. Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.}         & 20,800                  & 1941                      \\
RRUFF                                  & 1290                    & 2006                      \\
Crystallography Open Database (COD)    & 1063                    & 2003                      \\
PowBase                                & 169                     & 1999                      \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Linus Pauling File}:\cite{PaulingWeb} The Linus Pauling File is a largely commercial crystal structure database published and maintained by the Pauling File project \cite{villars2018}. It is currently distributed as Pearson Crystal data \cite{PearsonWeb} and the Materials Platform for Data Science (MPDS)\cite{MPDSWeb}. The database, first published in 2002, currently contains more than 534,000 crystal structures\cite{MPDSWeb} and 21,700 corresponding experimental powder diffraction patterns\cite{PearsonWeb}. 
This makes the Pauling file, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers. As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}. The MPDS is partially open, with the open MPDS data accessible through a web interface\cite{MPDSWeb}. Access to the full MPDS database can be purchased through a one-year license starting at \qty{2000}. \\

\textbf{Powder Diffraction File:} \cite{PDFWeb} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941\cite{GatesRector2019}. According to the ICDD the PDF5+, the latest release of the PDF as of November 2024, contains over a million materials with accompanying powder diffraction data. However, most of these powder diffraction patterns are simulated. After inquiring with the ICDD in April 2024 we were told that only 20,800 of the powder diffraction patterns in the PDF5+ stemmed from experiments. Of these 20,800 entries, 10,954 contain information about the atomic coordinates of the underlying structures. Since the PDF5+ lists the MPDS as a database source, there is likely a significant overlap in the experimental patterns found in the PDF5+ and those found in the Pauling file. As of November 2024, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265.00. However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published.\footnote{This information can be found in the product license agreement for the PDF5+ at \url{https://www.icdd.com/licensing-process/\#1528471154226-933e5cc6-8da7}.} \\

### 2024-12-04 chapters/04_our_dataset.tex
Added labeled column in dataset overview

\begin{table}[!htb]
\centering
\caption{Labeled opXRD database entries}
\label{tab:labeled}
\begin{tabular}{@{}ccc@{}} % Adjusted columns
\toprule
Institution & Research project & No. patterns\footnote{This number is subject to change} \\
\midrule
CNRS        & Extracted from the Crystallography open database & 895 \\
USC         & Study of CuNi and CuAl alloys & 338 \\
HKUST(GZ)   & Reference database for phase identification & 21 \\
EMPA        & Tin halide perovskites and combinatorial Zn-V-N library & 790 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!htb]
\centering
\caption{Unlabeled opXRD database entries}
\label{tab:unlabeled}
\begin{tabular}{@{}ccc@{}} % Adjusted columns
\toprule
Institution & Research project & No. patterns\footnote{This number is subject to change} \\
\midrule
INT         & Compilation of various research projects & 19780 \\
IKFT        & Compilation of various research projects & 64 \\
LBNL        & Time series recording of in-situ measurements & n/a \\
\bottomrule
\end{tabular}
\end{table}

@@

Experimental pXRD data was extracted from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}.

%We extracted experimental pXRD data from the Crystallography Open Database .

The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries. The data in the COD is placed in the public domain and licensed under the CC0 License. Of the entire COD database 5432 structures contained at least one tag from the {CIF\_POW} dictionary, i.e., a tag relating to powder diffraction studies. These 5432 structures only account for 1\% of the total COD database, but this is to be expected since most crystal structures are resolved from single-crystal diffraction. Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself. We could extract raw experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data. \\

>>>>

\begin{table}[!htb]
\centering
\caption{Merged opXRD database entries}
\label{tab:merged}
\begin{tabular}{@{}cccc@{}}
\toprule
Institution & Research Project & No. Patterns & Available labels \\ 
\midrule
CNRS        & Extracted from the COD         & 895          & Spg, Lattice, Unit cell, Comp   \\ 
HKUST(GZ)   & Phase identification reference database              & 21           & Spg, Lattice, Unit cell, Comp   \\ 
USC         & Study of CuNi and CuAl alloys                            & 338          & Spg, Lattice, Comp   \\ 
EMPA        & Tin halide perovskites, Zn-V-N library  & 790          & Spg, Comp   \\ 
INT         & Compilation of various projects                 & 19780        & None \\ 
IKFT        & Commercial catalysts, bulk materials                 & 64           & None \\ 
LBNL        & Time series recordings of Perovskites            & n/a          & None \\ 
\bottomrule
\end{tabular}
\end{table}

@@

Experimental pXRD data was extracted from the Crystallography Open Database (COD)\cite{Grazulis2009, Vaitkus2023}. The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, and metal-organic compounds and minerals, containing more than 500,000 entries. The data in the COD is placed in the public domain and licensed under the CC0 License. Of the entire COD database 5432 structures contained at least one tag from the {CIF\_POW} dictionary, i.e., a tag relating to powder diffraction studies. These 5432 structures only account for 1\% of the total COD database, but this is to be expected since most crystal structures are resolved from single-crystal diffraction. Of these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself. We could extract raw experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data. \\
